---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---


##Predicting health condition based on variables 
```{r}
data <- read.csv("ObesityDataSet.csv.xls", stringsAsFactors = TRUE)
```


#Since we have already done the correlations and summaries in Collab we already know what variables are most meaningful and which ones are not. 
```{r}
hist(data$Height)
#We can see most people are between 1.6 and 1.75 
```


```{r}
hist(data$Weight)
#We can see theres a second peak between 100 and 120Kg which are most likely going to be the people clasified with overweight or obesity. 
```

#Dropping not meaningful columns before startign the model building 
```{r}
data <- data[, !(names(data) %in% c('CALC', 'MTRANS', 'CAEC'))]
```



```{r}
head(data)
```

```{r}
library(dplyr)
```

```{r}
#Dropping all columns that are not going to be used for the model 

data <- data %>% select(-NCP)
data <- data %>% select(-SMOKE)
data <- data %>% select(-CH2O)
data <- data %>% select(-TUE)
data <- data %>% select(-SCC)

```


```{r}
#Changing the female to 1 and male to 0 to be able to build the model 
data$Gender <- as.character(data$Gender)  
data$Gender <- ifelse(data$Gender == "Female", "1", data$Gender)  
data$Gender <- ifelse(data$Gender == "Male", "0", data$Gender)  
 
```


```{r}
#chnaging the consumption of caloric food yes to 1 and no to 0 
data$FAVC <- as.character(data$FAVC)  
data$FAVC <- ifelse(data$FAVC == "yes", "1", data$FAVC)  
data$FAVC <- ifelse(data$FAVC == "no", "0", data$FAVC)  

```


```{r}
#CHanging the results to the following 
#'Insufficient_Weight' : 0
#'Normal_Weight' : 1
#'Overweight_Level_I' : 2
#'Overweight_Level_II' : 3
#'Obesity_Type_I' : 4
#'Obesity_Type_II' : 5
#'Obesity_Type_III' : 6

data$NObeyesdad <- as.character(data$NObeyesdad)  
data$NObeyesdad <- ifelse(data$NObeyesdad == "Insufficient_Weight", "0", data$NObeyesdad)  
data$NObeyesdad <- ifelse(data$NObeyesdad == "Normal_Weight", "1", data$NObeyesdad) 
data$NObeyesdad <- ifelse(data$NObeyesdad == "Overweight_Level_I", "2", data$NObeyesdad) 
data$NObeyesdad <- ifelse(data$NObeyesdad == "Overweight_Level_II", "3", data$NObeyesdad)
data$NObeyesdad <- ifelse(data$NObeyesdad == "Obesity_Type_I", "4", data$NObeyesdad)
data$NObeyesdad <- ifelse(data$NObeyesdad == "Obesity_Type_II", "5", data$NObeyesdad)
data$NObeyesdad <- ifelse(data$NObeyesdad == "Obesity_Type_III", "6", data$NObeyesdad)

#data$NObeyesdad <- factor(data$NObeyesdad) 
```


```{r}
#We want them as int not as factor thats why this is ## 
#data$Age <- factor(data$Age)
#data$Height <- factor(data$Height)
#data$Weight <- factor(data$Weight)
#data$FAF <- factor (data$FAF)
#data$FCVC <- factor(data$FCVC)
```

```{r}
#Making sure all of our variables are in the int form 
data$Gender <- as.integer(as.character(data$Gender))
data$Age <- as.integer(as.character(data$Age))
#data$Height <- as.integer(as.character(data$Height))
data$Weight <- as.integer(as.character(data$Weight))
#data$family_history_with_overweight <- as.integer(as.character(data$family_history_with_overweight))
data$FAVC <- as.integer(data$FAVC)
data$FCVC <- as.integer(as.character(data$FCVC))
data$FAF <- as.integer(as.character(data$FAF))


data$NObeyesdad <- as.numeric(as.factor(data$NObeyesdad))

```


```{r}
#install.packages("psych")
library(psych)
```

```{r}
cor(data)
```


#Start creating the model
#Here we are creatign the model and selecting which variables we want to use as dependent on the end estimator 
```{r}
ins_modeltwo <- lm(NObeyesdad ~ Gender + Age + Height + Weight + FAF  + FAVC + FCVC,
 data = data)

ins_modeltwo
```

```{r}
#Analyzing performance 
summary(ins_modeltwo)

#All variables seem to be significant since the value is greater than 2 as established in class. 

#The multiple r square being 0.9443 means that these variables are responsible at a 94.43% of in which category each person is put into  
```


```{r}
data$Age2 <- data$Age^2
```


```{r}
ins_modelfin <- lm(NObeyesdad ~ Gender + Age + Age2 + Height + Weight + FAF  + FAVC + FCVC, data = data)
```


```{r}
summary(ins_modelfin)
#We can see the adjustment didnt make that much of a difference. 
```



```{r}
#we are now going to use the model to make predictions 
data$pred <- predict(ins_modelfin, data)
cor(data$pred, data$NObeyesdad)


#The mdoel is 97.22% accurate which is fairly good 
```


#NOW LETS PREDICT 
#Gender = Female (1), Age(24), Height (1.91), Weight(70), FAF (2)
```{r}
predict(ins_modeltwo,
        data.frame(Gender = 1, Age = 24, Height = 1.91,
                   Weight = 70, FAF = 2 , FAVC = 1, FCVC = 2))
```


```{r}
predict(ins_modeltwo,
        data.frame(Gender = 0, Age = 24, Height = 1.89,
                   Weight = 90, FAF = 3 , FAVC = 1, FCVC = 3))
```

#this is a ramdom example that we got from the dataset and our model predicts this individual is normal weight just like it is on the data set 
```{r}
predict(ins_modeltwo,
        data.frame(Gender = 1, Age = 20, Height = 1.74,
                   Weight = 83, FAF = 2 , FAVC = 0, FCVC = 3))
#Considered normal weight
```



```{r}
predict(ins_modeltwo,
        data.frame(Gender = 0, Age = 23, Height = 1.66,
                   Weight = 60, FAF = 1 , FAVC = 0, FCVC = 3))
```





